The Talking Heads Experiment

### 2015-04-10 chap3/chap3.tex
cumulative changes to preface and chapter 1
updated labels, image widths, quotation marks, capitalisation, etc. for chapters 1-7

%\documentclass{../_combined/fcg-book}
\chapter{Perception}
Our minds deceive us. Intuitively we feel we see very clearly and 
unambiguously the objects in our environment and are able 
to categorise objects or actions unequivocally in terms of 
perceptually grounded distinctions. For example, when 
I look out of the office window at the inner 
courtyard garden, I can clearly see the plants, trees
and birds, as well as the walls, windows, and doors of 
the surrounding buildings. I can categorise their 
colours, sizes, and shapes and determine whether the 
leaves of the trees are moving with the wind. Because of 
this categorisation I can describe to another person what 
I see.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap3/figs/phototaxis}}
\caption{\label{phototaxis} 
Internal states of a robot's sensory and actuator channels on 
the y-axis and time in periods of 1/40 seconds on the x-axis.
The robot pushes against a box holding a light source.}
\end{figure}

@@

Similar behaviour systems and networks have
been experimented with 
for more complex tasks and it is actually the way that 
the Mars rover discussed in Chapter 1 works.\footnote{At 
the risk of simplifying, we can say that the early 
work in cybernetics (such as that of \cite{Braitenberg:1984}) 
has focused on this subsymbolic layer
and that {\it Artificial Intelligence}, as a research field
emerging in the late 1950's, has focused on the 
symbolic layer. Some researchers involved in a bottom up 
approach to Artificial Intelligence strongly refocused
on the subsymbolic layer. One of the most
vocal advocates of this position is \cite{Brooks:1992}.
Obviously we need the two, 
but bridging the gap is a non-trivial problem, sometimes known 
as the grounding problem. See: \cite{Harnad:1994}. }
However, the gap between the continuous dynamics of sensori-motor
intelligence and cognition remains unbridged. 
It is possible for {\itshape us} to see 
structure in the data but this structure is not perceived
nor used for control by the robot itself. 
The robot does not segment nor categorise
reality. It does not `know' that it is moving left or right 
and therefore cannot communicate this information to another robot. 
All processing remains at the analog continuous
level. However because it remains at this subsymbolic level, it is doubtful
whether we can ever hope to bootstrap cognitive
intelligence simply by adding more of these dynamical 
systems.\footnote{
\cite{Marr:1982} remains a classical reference outlining
the features that can be extracted, the
principled algorithms for doing it, and possible 
neurophysiological models. These are the main research
topics that still dominate research in vision. See 
\cite{Ullman:1996} for an overview.}
The difference between behavioural intelligence 
and cognitive intelligence resides in an additional 
layer of processing which is no longer continuous and analog 
but discrete and symbolic. How this second symbolic 
layer could be formed but at the same time remain 
grounded in the analog sensori-motor layer is one 
of the main research topics addressed in this book.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap3/figs/sslang}}
\label{screenshot}
\caption{ Example of a scene captured by the 
camera in \figref{f:plate7}, containing a puppet, a
wooden house, and a horse. Segmentation is based on aggregating grayscale
patches, i.e. areas in the image that are lighter or darker than the 
general background. Bounding boxes have been drawn around the segments. 
Note that there can be bounding boxes within bounding boxes if an object 
forms part of a larger object. 
\label{f:boundbox}}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap3/figs/image1}}
\caption{ Left view shows an image as captured by a Talking
Heads camera. Middle view shows the result of 
segmentation based on edge detection. Right view
shows the integration of segmentation by color
and by edge detection.}
\label{f:plate11}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap3/figs/diff-percept}}
\caption{ Top (left), image captured by the hearer. 
Bottom (left), image captured by the speaker. Because
both have a different view on the scene, the 
images diverge and consequently 
the segmentation (bottom and top right) diverges as well, in particular the yellow 
circle is not perceived by the speaker. In contrast the speaker perceives a rectangle
(bottom left corner) and has chosen this as the topic. The rectangle is not perceived by the 
hearer, so the game has no chance to be successful.}
\label{f:plate12}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap3/figs/sieve}}
\caption{ The different layers of cognitive processing act 
like a sieve. Inputs flow into the layer, where they are processed
to generate hypotheses of which the best ones are transmitted
to the next layer. Every layer can operate in both directions, 
so that constraints can flow from top to bottom and from bottom
to top. \label{f:sieve}}
\end{figure}

@@

Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (HPOS) and vertical position (VPOS) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \ref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the BABEL environment. 
}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/recscene}}
\caption{ \label{rect1b} A scene with three objects. The
sensory values on the grayscale channel are the most salient 
and will therefore be chosen preferentially for categorisation
and verbalisation.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10}
\end{center}
\caption{Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being to the right and so the same word 
`gofubo' has been used to refer to them. }
\label{f:plate-10}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap3/figs/scaling}}
\caption{\label{scaling} Processing converts data on 
a sensory channel into more usable data for later categorisation processes.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap3/figs/square}}
\caption{\label{square} The general problem of 
production is broken up into three subproblems: perception, 
conceptualisation and verbalisation.}
\end{figure}

@@

The methodology of breaking a problem into its subproblems goes quite 
a distance, but is not without danger. The processes relating
perception to language cannot be modular for reasons already mentioned. 
Each layer receives inputs which are not completely reliable and 
generates a set of possible suggestions rather than a single
`correct' solution. Constraints have to flow
from top to bottom because there is simply not enough reliable 
information to solve the problem with a straightforward sequential 
decision process. In addition, every layer is constantly adapting 
itself to the surrounding information context. New categories 
may develop, new words need to be learned, new sensory 
channels may even emerge. So, constructing a global system is going 
to be more than simply putting together its parts. There are complex
behaviours that will only become visible when the appropriate non-modular
couplings are put in place, and this is notoriously difficult to 
do and study.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/scene9}}
\caption{ \label{geom3} An example of the 
computer-generated scenes from the GEOM world.
Each shape is labeled for further reference.}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom} Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled} Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}

>>>>

%\documentclass{../_combined/fcg-book}
\chapter{Perception} \label{chap:3}
Our minds deceive us. Intuitively we feel we see very clearly and 
unambiguously the objects in our environment and are able 
to categorise objects or actions unequivocally in terms of 
perceptually grounded distinctions. For example, when 
I look out of the office window at the inner 
courtyard garden, I can clearly see the plants, trees
and birds, as well as the walls, windows, and doors of 
the surrounding buildings. I can categorise their 
colours, sizes, and shapes and determine whether the 
leaves of the trees are moving with the wind. Because of 
this categorisation I can describe to another person what 
I see.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap3/figs/phototaxis.pdf}}
\caption{\label{phototaxis}Internal states of a robot's sensory and actuator channels on 
the y-axis and time in periods of 1/40 seconds on the x-axis.
The robot pushes against a box holding a light source.}
\end{figure}

@@

Similar behaviour systems and networks have
been experimented with 
for more complex tasks and it is actually the way that 
the Mars rover discussed in Chapter 1 works.\footnote{At 
the risk of simplifying, we can say that the early 
work in cybernetics (such as that of \cite{Braitenberg:1984}) 
has focused on this subsymbolic layer
and that {\it Artificial Intelligence}, as a research field
emerging in the late 1950's, has focused on the 
symbolic layer. Some researchers involved in a bottom up 
approach to Artificial Intelligence strongly refocused
on the subsymbolic layer. One of the most
vocal advocates of this position is \cite{Brooks:1992}.
Obviously we need the two, 
but bridging the gap is a non-trivial problem, sometimes known 
as the grounding problem. See: \cite{Harnad:1994}. }
However, the gap between the continuous dynamics of sensori-motor
intelligence and cognition remains unbridged. 
It is possible for {\itshape us} to see 
structure in the data but this structure is not perceived
nor used for control by the robot itself. 
The robot does not segment nor categorise
reality. It does not ``know'' that it is moving left or right 
and therefore cannot communicate this information to another robot. 
All processing remains at the analog continuous
level. However because it remains at this subsymbolic level, it is doubtful
whether we can ever hope to bootstrap cognitive
intelligence simply by adding more of these dynamical 
systems.\footnote{
\cite{Marr:1982} remains a classical reference outlining
the features that can be extracted, the
principled algorithms for doing it, and possible 
neurophysiological models. These are the main research
topics that still dominate research in vision. See 
\cite{Ullman:1996} for an overview.}
The difference between behavioural intelligence 
and cognitive intelligence resides in an additional 
layer of processing which is no longer continuous and analog 
but discrete and symbolic. How this second symbolic 
layer could be formed but at the same time remain 
grounded in the analog sensori-motor layer is one 
of the main research topics addressed in this book.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap3/figs/sslang.pdf}}
\label{screenshot}
\caption{Example of a scene captured by the 
camera in \figref{f:plate7}, containing a puppet, a
wooden house, and a horse. Segmentation is based on aggregating grayscale
patches, i.e. areas in the image that are lighter or darker than the 
general background. Bounding boxes have been drawn around the segments. 
Note that there can be bounding boxes within bounding boxes if an object 
forms part of a larger object. 
\label{f:boundbox}}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap3/figs/image1.pdf}}
\caption{Left view shows an image as captured by a Talking
Heads camera. Middle view shows the result of 
segmentation based on edge detection. Right view
shows the integration of segmentation by color
and by edge detection.}
\label{f:plate11}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap3/figs/diff-percept.pdf}}
\caption{Top (left), image captured by the hearer. 
Bottom (left), image captured by the speaker. Because
both have a different view on the scene, the 
images diverge and consequently 
the segmentation (bottom and top right) diverges as well, in particular the yellow 
circle is not perceived by the speaker. In contrast the speaker perceives a rectangle
(bottom left corner) and has chosen this as the topic. The rectangle is not perceived by the 
hearer, so the game has no chance to be successful.}
\label{f:plate12}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap3/figs/sieve.pdf}}
\caption{The different layers of cognitive processing act 
like a sieve. Inputs flow into the layer, where they are processed
to generate hypotheses of which the best ones are transmitted
to the next layer. Every layer can operate in both directions, 
so that constraints can flow from top to bottom and from bottom
to top. \label{f:sieve}}
\end{figure}

@@

Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (HPOS) and vertical position (VPOS) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \ref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the \textsc{babel} environment. 
}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/recscene.pdf}}
\caption{\label{rect1b}A scene with three objects. The
sensory values on the grayscale channel are the most salient 
and will therefore be chosen preferentially for categorisation
and verbalisation.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10.pdf}
\end{center}
\caption{Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being to the right and so the same word 
`gofubo' has been used to refer to them. }
\label{f:plate-10}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap3/figs/scaling.pdf}}
\caption{\label{scaling}Processing converts data on 
a sensory channel into more usable data for later categorisation processes.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap3/figs/square.pdf}}
\caption{\label{square}The general problem of 
production is broken up into three subproblems: perception, 
conceptualisation and verbalisation.}
\end{figure}

@@

The methodology of breaking a problem into its subproblems goes quite 
a distance, but is not without danger. The processes relating
perception to language cannot be modular for reasons already mentioned. 
Each layer receives inputs which are not completely reliable and 
generates a set of possible suggestions rather than a single
``correct'' solution. Constraints have to flow
from top to bottom because there is simply not enough reliable 
information to solve the problem with a straightforward sequential 
decision process. In addition, every layer is constantly adapting 
itself to the surrounding information context. New categories 
may develop, new words need to be learned, new sensory 
channels may even emerge. So, constructing a global system is going 
to be more than simply putting together its parts. There are complex
behaviours that will only become visible when the appropriate non-modular
couplings are put in place, and this is notoriously difficult to 
do and study.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/scene9.pdf}}
\caption{\label{geom3}An example of the 
computer-generated scenes from the GEOM world.
Each shape is labeled for further reference.}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom}Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled}Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}
